---
title: "charlie_slides"
author: "Charlie Edelson"
date: "11/30/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```


## Rate vs Death Score

```{r, waring=FALSE, message=FALSE}
setwd("/Users/edelsonc/Desktop/Data_Science/Group_EDA/EDA_Project_2/Data_concise")
library(tidyverse)
na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
df_c <- read_csv("edelsonc_merged.csv")


ggplot(data=df_c, aes(x=Overall.Score, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm')

lm_rd <- lm(Deaths.Score ~ Overall.Score, data=df_c)
summary(lm_rd)

# obesity
ggplot(data=df_c, aes(x=Percent.Population.Obese, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm')

lm_od <- lm(Deaths.Score ~ Percent.Population.Obese, data = df_c)
summary(lm_rd)

# median income
ggplot(data=df_c, aes(x=Median.Income, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm')

lm_md <- lm(Deaths.Score ~ Median.Income, data=df_c) 
summary(lm_md)
```

## Granularity
Data Comes in Three Levels

- National
- State
- Hospital

Aggregated into County

<ul>
<li>National level too fuzzy, hospital too fine</li>
<li>Averaging behavior </li>
<ul>
  <li>Halved Number Rows (4800 $\rightarrow$ 2000)</li>
</ul>
</ul>

## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/edelsonc/Desktop/Data_Science/Group_EDA/EDA_Project_2/Data_concise")
library(tidyverse)
na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
df_c <- read_csv("edelsonc_merged.csv")


ggplot(data=df_c, aes(x=Overall.Score, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Average County Overall Rating") + ylab("Death Score")
```

***

```{r, warning=FALSE, message=FALSE}
lm_rd <- lm(Deaths.Score ~ Overall.Score, data=df_c)
summary(lm_rd)
```