---
title: "charlie_slides"
author: "Charlie Edelson"
date: "11/30/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

## Granularity
Data Comes in Three Levels

- National
- State
- Hospital

## Granularity
Data Comes in Three Levels

- National
- State
- Hospital

Aggregated into County

<ul>
<li>National level too fuzzy, hospital too fine</li>
<li>Averaging behavior </li>
<ul>
  <li>Halved Number Rows (4800 $\rightarrow$ 2000)</li>
</ul>
</ul>

## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.1}
setwd("/Users/edelsonc/Desktop/Data_Science/Group_EDA/EDA_Project_2/Data_concise")
library(tidyverse)
na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
df_c <- read_csv("edelsonc_merged.csv")


ggplot(data=df_c, aes(x=Overall.Score, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Average County Overall Rating") + ylab("Death Score")
```


## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.1}
setwd("/Users/edelsonc/Desktop/Data_Science/Group_EDA/EDA_Project_2/Data_concise")
library(tidyverse)
na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
df_c <- read_csv("edelsonc_merged.csv")


ggplot(data=df_c, aes(x=Overall.Score, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Average County Overall Rating") + ylab("Death Score")
```

- Significant Intercept (t = 62.6, p < 2 x 10<sup>-16</sup>)
- Significant Coefficient (t = -5.034, p< 6 x 10<sup>-7</sup>)
- Bad R<sup>2</sup> (0.03); Strong F-stat (F = 25.42, p < 10<sup>-7</sup>)

## Recurring Theme
```{r, warnings=FALSE, message= FALSE, echo=FALSE, fig.height=3.3}
library(gridExtra)

p1 <- ggplot(data=df_c, aes(x=Percent.Population.Obese, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Percent Obesity") + ylab("Death Score") + ggtitle("Obesity Percentage vs Death Score")

p2 <- ggplot(data=df_c, aes(x=Median.Income, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm')+ xlab("Median Income") + ylab("Death Score") + ggtitle("Median Income vs Death Score")

grid.arrange(p1, p2, ncol=2)
```

- Significant Intercepts and Significant Coefficients
- Poor R<sup>2</sup> ( < 0.08) but Strong F-Statistic


