---
title: "Project 2 Hospitals and MI's"
author: "Matthew McAvoy"
date: "November 30, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r include=FALSE}
library(tidyverse)
library(stringr)
library(ggrepel)

setwd("C:/Users/homur/OneDrive/New College/EDA/Project2")
data <- read.csv("Data_concise/merged_data.csv")
```

## Overall Score
<span style="color:midnightblue">$\uparrow$</span> Overall.Score - Summarizes more than 60 measures reported on Hospital Compare into a single rating. 
Seven groups:

- Mortality
- Safety of care
- Readmission
- Patient experience
- Effectiveness of care
- Timeliness of care
- Efficient use of medical imaging

## Variables

Location - US Counties.

<span style="color:deepskyblue">$\downarrow$</span> Compl.Score - Measures were complications like 'Blood stream infection after surgery' and 'Serious blood clots after surgery'. 

<span style="color:darkcyan">$\uparrow$</span> Quests.Score - A questionnaire scored a hospital on a star system of 1-5. Some questions were 'room was "always clean"', and 'Doctors "always" communicated well'.

<span style="color:royalblue">$\downarrow$</span> Infects.score - Includes blood stream infections, UTI's, infections after surgery, and MRSA and C.Diff (Clostridium difficile) infections.

<span style="color:orchid">$\uparrow$</span> Spends.Score - Medicare Hospital Spending per patient. (Larger better?)

## Variables
<span style="color:steelblue">$\downarrow$</span> Images.score - Such imaging as Cat scans of abdomens, MRI's of spines, and mammography follow-up rates. Up or down?

<span style="color:mediumslateblue">$\downarrow$</span> Values.Payment - Assess Medicare Part A and Part B payments for services. Are price-standardized and risk-adjusted. Purpose is to see whether payments for treatments at a particular hospital is less than or no different than the national average. Retained only payments for acute MI's.

<span style="color:blue">$\downarrow$</span> Deaths.Score - Retained measure Acute MI 30 day mortality rate. 

<span style="color:cyan">$\downarrow$</span> Timely.Score -  Typically the number of minutes waited for treating MI's by use of fibrinolytic therapy, transfer to another facility for better treatent, Aspirin on arrival, or time to ECG.


## Relationship of overall score to groups, comparing top and bottom 20 counties
```{r}
top_score <- data %>% arrange(desc(Overall.Score)) %>% slice(1:20)
bottom_score <- data %>% arrange(Overall.Score) %>% slice(1:20)

bottom_top <- bind_rows(top_score %>% mutate(Rank = "Top"), bottom_score %>% mutate(Rank = "Bottom"))

```

## Rank to Overall Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Overall.Score, color=Rank))
```

## Rank to Questionnaire Stars
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Quests.Stars, color=Rank))
```

## Rank to Complications
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Compl.Score, color=Rank))
```

## Rank to Infections Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Infects.Score, color=Rank))
```

## Rank to Spending Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Spends.Score, color=Rank))
```

## Rank to Payment Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Values.Payment, color=Rank))
```

## Rank to Deaths Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Deaths.Score, color=Rank))
```

## Rank to Timely Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Timely.Score, color=Rank))
```

## Boxplot summary
See general trend of overall top ranked counties have lower scores (lower is better) compared to worst ranked counties.

## State level
Aggregated on States to find States with the best overall score and the lowest Death score.
```{r}
state_data <- data %>% mutate(State = str_extract(Location, "\\s..$"))
state_score <- state_data %>% group_by(State) %>% summarise(State.Overall.Score = mean(Overall.Score), State.Deaths.Score = mean(Deaths.Score))
```

## State level by Overall Score
```{r}
ggplot(data = state_score, aes(x=State, y=State.Overall.Score)) + 
		geom_point(shape = 1, size = 2.5) + geom_text_repel(aes(label=State), size = 4)
```

## State level by Death Score
```{r}
ggplot(data = state_score, aes(x=State, y=State.Deaths.Score)) + 
		geom_point(shape = 1, size = 2.5) + geom_text_repel(aes(label=State), size = 4)
```

## Summary of State level
North Dakota has a very high death score and Rhode Island, Utah, and Delaware have a small death score.


