---
title: "Hospitals and MI's"
author: "Caleb Dowdy, Charlie Edelson, Matthew McAvoy, Eric Voorhis"
date: "December 1, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r include=FALSE}
library(tidyverse)
library(stringr)
library(ggrepel)
library(datasets)

setwd("C:/Users/homur/OneDrive/New College/EDA/Project2")
data1 <- read.csv("Data_concise/merged_data.csv")
data <- read.csv("Data_concise/edelsonc_merged.csv")

na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
quests <- read_csv("Data/HCAHPS - Hospital.csv", na=na_list)


completeness <- function(dataframe) {
  df <- bind_rows(
    lapply(dataframe, function(x) {
      data.frame(nmiss=sum(is.na(x)),
                 n=length(x),
                 completeness= 1 - (sum(is.na(x))/length(x)))
    })
  )
  cbind(colname=colnames(dataframe), df)
}

na_list <- c("", " ", "Not Available", "NA", "NAN", "Not Applicable")

process_csv <- function(file_name) {
  df <- read.csv(file_name, na.strings=na_list)
  df$State <- as.character(df$State)
  fifty <- which(df$State %in% state.abb)
  datareturn <- df[fifty,]
  return(datareturn)
}

mortality_types <- c("Acute Myocardial Infarction (AMI) 30-Day Mortality Rate",
                     "Heart failure (HF) 30-Day Mortality Rate")

df_surveys <- process_csv("Data/HCAHPS - Hospital.csv")
df_death <- process_csv("Data/Readmissions and Deaths - Hospital.csv")
df_ratings <- process_csv("Data/Hospital General Information.csv")
df_payments <- process_csv("Data/Payment and Value of Care - Hospital.csv")
df_timely <- process_csv("Data/Timely and Effective Care - Hospital.csv")

plotCompleteness <- function(data_frame, title) {
  completion_df <- completeness(data_frame)
  ggplot(completion_df, aes(x=reorder(factor(colname), completeness), y=completeness)) + 
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
}

df_death_subset_ami <- df_death %>%
  filter(Measure.Name == "Acute Myocardial Infarction (AMI) 30-Day Mortality Rate")

df_death_subset_hf <- df_death %>%
  filter(Measure.Name == "Heart failure (HF) 30-Day Mortality Rate")

death_completeness <- completeness(df_death)
death_completeness_ami <- completeness(df_death_subset_ami)
death_completeness_hf <- completeness(df_death_subset_hf)

completeness_ami_by_state <- df_death_subset_ami %>%
  group_by(State) %>%
  summarise(nmiss=sum(is.na(Score)),
            n=length(Score),
            completeness=1-(sum(is.na(Score))/length(Score)))

completeness_by_state_all <- df_death %>%
  group_by(State) %>%
  summarise(nmiss=sum(is.na(Score)),
            n=length(Score),
            completeness=1-(sum(is.na(Score))/length(Score)))
  
completeness_ami_by_state$subset_type <- "Myocardial Infarction"
completeness_by_state_all$subset_type <- "All"
```
# Introduction

## Question of Investigation
Where is a good or bad place to have a heart attack?

## Statement of Thesis
Higher rated counties are better places to have heart attacks

Significant finding - Found relationship between Death Score and Overall Rating

## Data 

- Provided by Hospital Compare website
    - Consists of a variety of metrics related to the performance measures 
- Over 4,000 Medicare-certified hospitals are included in their databases
- Medicare billing codes 
- Center of Medicare & Medicaid Services
- "Wide range of readers"

## Granularity
Data Comes in Three Levels

- National
- State
- Hospital

## Granularity
Data Comes in Three Levels

- National
- State
- Hospital

Aggregated into County

<ul>
<li>National level too fuzzy, hospital too fine</li>
<li>Averaging behavior </li>
<ul>
  <li>Halved Number Rows (4800 $\rightarrow$ 2000)</li>
</ul>
</ul>

## Completeness

Missing values show up in several forms:

* ""
* " "
* "Not Available"
* "NA"
* "NAN"
* "Not Applicable"

## Completeness of flat files

```{r }
plotCompleteness(df_surveys, "HCAHPS Completeness")
```

## Completeness of flat files
```{r}
death_completeness <- completeness(df_death)
death_completeness <- death_completeness[order(death_completeness$completeness),]
  
plot_df <- death_completeness

plot_df$colname <- factor(plot_df$colname, as.character(plot_df$colname))

ggplot(plot_df, aes(x=factor(colname), y=completeness)) + 
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    xlab("Column") + ylab("Completeness") + ggtitle("Readmission and Death Completeness")

```

## Completeness of flat files
```{r}
plot_df <- completeness(df_death_subset_ami)

target <- death_completeness$colname
plot_df <- plot_df[match(target, plot_df$colname),]

plot_df$colname <- factor(plot_df$colname, as.character(plot_df$colname))

ggplot(plot_df, aes(x=factor(colname), y=completeness)) + 
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    xlab("Column") + ylab("Completeness") + ggtitle("Readmission and Death Completeness - (AMI) Subset")
```

## Completeness by state
```{r warning=FALSE, fig.width = 9, fig.height = 4.5}
states_map <- map_data("state")

state_frame <- data.frame(State=state.abb, state=state.name)
state_frame$state <- str_to_lower(state_frame$state)

df <- inner_join(completeness_ami_by_state, state_frame, by="State")

p <- ggplot(df, aes(map_id=state)) +
  geom_map(aes(fill = completeness), map = states_map, color = "black", size = 0.25) +
  expand_limits(x = states_map$long, y = states_map$lat) + xlab("Long") + ylab("Lat") +
  ggtitle("Acute Myocardial Infarction (AMI) 30-Day Mortality Rate Completeness")

p + coord_map()
```

# Quantifying "Worst"

## Who is actually included? 
An **Index Admission** is the hospitiliztion to which the mortality outcome is attributed and includes patients that meet the following criteria: 

- Aged 65 or over
- Enrolled in Original Medicare for 12 months prior to admission
- Consistent or known vital status with clear demographic (age and gender data)
- **Not** discharged against medical advice
- **Not** transferred from another acute care facility
- **Not** enrolled in Medicare hospice program any time in the 12 months prior to admission
- **Many more special cases** ask us about it after...

## What is the Death Score? 

`RSMR` $= \frac{"predicted"}{"expected"}\times National \ NMR$ 

- `predicted` : Number of deaths within 30 days based on the hospitals performance for a given patient mix 
- `expected` : Number of deaths expected based on the national average for hospital performance with the same patient mix 
- `NMR`: National mortality rate

## Overall Score
<span style="color:midnightblue">$\uparrow$</span> Overall.Score - Summarizes more than 60 measures reported on Hospital Compare into a single rating. 
Seven groups:

- Mortality
- Safety of care
- Readmission
- Patient experience
- Effectiveness of care
- Timeliness of care
- Efficient use of medical imaging

## Variables

Location - US Counties.

<span style="color:deepskyblue">$\downarrow$</span> Compl.Score - Measures were complications like 'Blood stream infection after surgery' and 'Serious blood clots after surgery'. 

<span style="color:darkcyan">$\uparrow$</span> Quests.Score - A questionnaire scored a hospital on a star system of 1-5. Some questions were 'room was "always clean"', and 'Doctors "always" communicated well'.

<span style="color:royalblue">$\downarrow$</span> Infects.score - Includes blood stream infections, UTI's, infections after surgery, and MRSA and C.Diff (Clostridium difficile) infections.

<span style="color:orchid">$\downarrow$</span> Spends.Score - Standardized Medicare Hospital Spending per patient. Less than 1 means Medicare is cheaper at this location, can be regarded as good.

## Variables
<span style="color:steelblue">$\downarrow$</span> Images.score - Such imaging as Cat scans of abdomens, MRI's of spines, and mammography follow-up rates. Up or down?

<span style="color:mediumslateblue">$\downarrow$</span> Values.Payment - Assess Medicare Part A and Part B payments for services. Are price-standardized and risk-adjusted. Purpose is to see whether payments for treatments at a particular hospital is less than or no different than the national average. Retained only payments for acute MI's.

<span style="color:blue">$\downarrow$</span> Deaths.Score - Retained measure Acute MI 30 day mortality rate. 

<span style="color:cyan">$\downarrow$</span> Timely.Score -  Typically the number of minutes waited for treating MI's by use of fibrinolytic therapy, transfer to another facility for better treatent, Aspirin on arrival, or time to ECG.

## Relationship of overall score to groups, comparing top and bottom 20 counties
```{r}
top_score <- data %>% arrange(desc(Overall.Score)) %>% slice(1:20)
bottom_score <- data %>% arrange(Overall.Score) %>% slice(1:20)

bottom_top <- bind_rows(top_score %>% mutate(Rank = "Top"), bottom_score %>% mutate(Rank = "Bottom"))

```

## Rank to Overall Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Overall.Score, color=Rank)) + ylab("Stars") + theme_minimal()
```

## Rank to Questionnaire Stars
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Quests.Stars, color=Rank)) + ylab("Stars") + theme_classic()
```

## Rank to Complications
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Compl.Score, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Infections Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Infects.Score, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Spending Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Spends.Score, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Payment Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Values.Payment, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Deaths Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Deaths.Score, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Timely Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Timely.Score, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Obesity Score
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Percent.Population.Obese, color=Rank)) + ylab("Score") + theme_classic()
```

## Rank to Median Income
```{r}
ggplot(data = bottom_top, aes(x=Rank)) + geom_boxplot(aes(y=Median.Income, color=Rank)) + ylab("Income") + theme_classic()
```

## Boxplot summary
See general trend of overall top ranked counties have lower scores (lower is better) compared to worst ranked counties.

## State level
Aggregated on States to find States with the best overall score and the lowest Death score.
```{r}
state_data <- data %>% mutate(State = str_extract(Location, "\\s..$"))
state_score <- state_data %>% group_by(State) %>% summarise(State.Overall.Score = mean(Overall.Score), State.Deaths.Score = mean(Deaths.Score))
# old
color_green <- state_score %>% filter(str_detect(State, 'UT|RI|DE'))
color_red <- state_score %>% filter(str_detect(State, 'AK|MS|WV|ND'))
# new
color_green <- state_score %>% filter(str_detect(State, 'ID|MT|DE|IN|MN'))
color_red <- state_score %>% filter(str_detect(State, 'NY|NV|MS|NJ|FL'))
```

## County level by Overall Score
```{r}
top_five <- data %>% arrange(desc(Overall.Score)) %>% slice(1:5)
bottom_five <- data %>% arrange(Overall.Score) %>% slice(1:5)
ggplot(data = data, aes(x=Location, y=Overall.Score)) + 
		geom_point(shape = 1, size = 2.5) + 
    geom_text_repel(data=top_five, aes(label=Location), color="forestgreen", size = 4) + 
    geom_text_repel(data=bottom_five, aes(label=Location), color="red", size = 4) + theme_minimal() + xlab("County") + ylab("Overall Score") + theme(axis.text.x=element_blank())
```

## County level by Death Score
```{r}
worst_five <- data %>% arrange(desc(Deaths.Score)) %>% slice(1:5)
best_five <- data %>% arrange(Deaths.Score) %>% slice(1:5)
ggplot(data = data, aes(x=Location, y=Deaths.Score)) + 
		geom_point(shape = 1, size = 2.5) + 
    geom_text_repel(data=best_five, aes(label=Location), color="forestgreen", size = 4)+ 
    geom_text_repel(data=worst_five, aes(label=Location), color="red", size = 4) + theme_minimal() + xlab("County") + ylab("Death Score") +  theme(axis.text.x=element_blank()) 
```

## State level by Overall Score
```{r}
ggplot(data = state_score, aes(x=State, y=State.Overall.Score)) + 
		geom_point(shape = 1, size = 2.5) + geom_text_repel(aes(label=State), size = 4) + 
		geom_point(data=color_green, aes(x=State, y=State.Overall.Score), color="forestgreen", shape= 16, size = 2.5) + 
		geom_point(data=color_red, aes(x=State, y=State.Overall.Score), color="red", shape= 16, size = 2.5) + theme_minimal() + xlab("State") + ylab("Overall Score") + theme(axis.text.x=element_blank()) 

```

## State level by Death Score
```{r}
worst_five <- state_score %>% arrange(desc(State.Deaths.Score)) %>% slice(1:5)
best_five <- state_score %>% arrange(State.Deaths.Score) %>% slice(1:5)
ggplot(data = state_score, aes(x=State, y=State.Deaths.Score)) + 
		geom_point(shape = 1, size = 2.5) + geom_text_repel(aes(label=State), size = 4) + 
		geom_point(data=best_five, aes(x=State, y=State.Deaths.Score), color="forestgreen", shape= 16, size = 2.5) + 
		geom_point(data=worst_five, aes(x=State, y=State.Deaths.Score), color="red", shape= 16, size = 2.5) + theme_minimal() + xlab("State") + ylab("Death Score") + 
		theme(axis.text.x=element_blank())
```


```{r include = FALSE}
quests_q2 <- quests %>% filter(str_detect(`HCAHPS Measure ID`,'_SCORE')) %>% 
		mutate(Location = paste(`County Name`, State)) %>% 
		rename(Score = `HCAHPS Linear Mean Value`) %>%
		rename(Question = `HCAHPS Answer Description`) %>%
		select(Location, Question, Score) %>% filter(!is.na(Score)) %>%
		mutate_each(funs(as.numeric), one_of(c("Score"))) %>% 
		group_by(Question) %>% summarise(Quests.Score = mean(Score))
```

## National Questionnaire Performance 
```{r}
ggplot(data = quests_q2, aes(x=Question, y=Quests.Score)) + 
		geom_point(shape = 1, size = 2.5) + geom_text_repel(aes(label=Question), size = 4) + 
  theme_minimal() + xlab("Questions") + ylab("Score") + theme(axis.text.x=element_blank()) 
```

## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.1}
library(tidyverse)
na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
df_c <- read_csv("Data_concise/edelsonc_merged.csv")


ggplot(data=df_c, aes(x=Overall.Score, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Average County Overall Rating") + ylab("Death Score")
```


## Effect of Overall Rating on Death Score

- Nice if Overall Rating was related to Death Score

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.1}
library(tidyverse)
na_list <- c("", " ", "Not Available", "NA", "NAN", "NaN")
df_c <- read_csv("Data_concise/edelsonc_merged.csv")


ggplot(data=df_c, aes(x=Overall.Score, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Average County Overall Rating") + ylab("Death Score")
```

- Significant Intercept (t = 62.6, p < 2 x 10<sup>-16</sup>)
- Significant Coefficient (t = -5.034, p< 6 x 10<sup>-7</sup>)
- Bad R<sup>2</sup> (0.03); Strong F-stat (F = 25.42, p < 10<sup>-7</sup>)

## Recurring Theme
```{r, warning=FALSE, message= FALSE, echo=FALSE, fig.height=3.3}
library(gridExtra)

p1 <- ggplot(data=df_c, aes(x=Percent.Population.Obese, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm') + xlab("Percent Obesity") + ylab("Death Score") + ggtitle("Obesity Percentage vs Death Score")

p2 <- ggplot(data=df_c, aes(x=Median.Income, y=Deaths.Score)) + geom_point() + geom_smooth(method='lm')+ xlab("Median Income") + ylab("Death Score") + ggtitle("Median Income vs Death Score")

grid.arrange(p1, p2, ncol=2)
```

- Significant Intercepts and Significant Coefficients
- Poor R<sup>2</sup> ( < 0.08) but Strong F-Statistic


## Conclusion
With an overall score already computed, better counties are better places to have heart attacks, supporting our initial hypothesis


## Thoughts and Further Investigations
- Broader data set
- More variables, preferably not already standardized

- Closing remarks

